### 2015-10-22 writing/gist/model1_fig.tex

\begin{figure*}[t]
\begin{subfigure}[t]{\textwidth}
\centering
\scalebox{0.6}{
 \begin{tikzpicture}
    \def\dx{21pt}
    \def\dy{11pt}
    \def\sy{13*\dy}
    \def\oxb{8*\dx}
    \def\by{1*\dy}
    \def\ox{0*\oxb}

    \tikzstyle{word}=[text width=35mm,align=center,text height=2mm]    
    \tikzstyle{tracker}=[fill=red!40,text width=15mm,align=center,text height=2mm]
    \tikzstyle{softmax}=[fill=blue!40,text width=15mm,align=center,text height=2mm]
    \tikzstyle{comp}=[fill=green!40,text width=20mm,align=center,text height=2mm]
    \tikzstyle{result}=[line width=1pt,draw=black,text width=15mm,align=center,text height=2mm]    
    \tikzstyle{sbox}=[line width=1pt,draw=black,text width=25mm,align=center,text height=10mm]
    \tikzstyle{bbox}=[line width=1pt,draw=black,text width=25mm,align=center,text height=6.5mm]
    \tikzstyle{focus1}=[fill=yellow!80,opacity=0.5,text width=25mm,align=center,text height=2mm]
    \tikzstyle{focus2}=[fill=yellow!80,opacity=0.5,text width=25mm,align=center,text height=5.5mm]

    \node[focus1] (0bb) at  (\ox+0*\dx,2*\dy) {};
    \node[word]  (0b3) at (\ox+0*\dx,\by-1*\dy) {};
    \node[word]  (0b2) at (\ox+0*\dx,\by+0*\dy) {down};
    \node[word]  (0b1) at (\ox+0*\dx,\by+1*\dy) {sat};
    \node[bbox] (0bb) at  (\ox+0*\dx,\by+0.5*\dy) {};

    \node[focus2] (0sb) at  (\ox+0*\dx,\sy-0.5*\dy) {};
    \node[word]  (0s1) at (\ox+0*\dx,\sy-1*\dy) {cat};
    \node[word]  (0s2) at (\ox+0*\dx,\sy+0*\dy) {the};
    \node[word]  (0s3) at (\ox+0*\dx,\sy+1*\dy) {};
    \node[sbox] (0sb) at  (\ox+0*\dx,\sy+0*\dy) {};

    \node[comp] (0c) at  (\ox+0.5*\oxb,\sy-1.5*\dy) {composition};

    \node[tracker] (0t) at  (\ox+0*\dx,5*\dy) {tracking};
    \node[softmax] (0sm) at  (\ox+3*\dx,7*\dy) {$\sigma$};
    \node[result] (0so) at  (\ox+3*\dx,9*\dy) {\reduce};

    \def\ox{1*\oxb}

    \node[focus1] (1bb) at  (\ox+0*\dx,2*\dy) {};
    \node[word]  (1b3) at (\ox+0*\dx,\by-1*\dy) {};
    \node[word]  (1b2) at (\ox+0*\dx,\by+0*\dy) {down};
    \node[word]  (1b1) at (\ox+0*\dx,\by+1*\dy) {sat};
    \node[bbox] (1bb) at  (\ox+0*\dx,\by+0.5*\dy) {};

    \node[focus2] (1sb) at  (\ox+0*\dx,\sy-0.5*\dy) {};
    \node[word]  (1s1) at (\ox+0*\dx,\sy-1*\dy) {the cat};
    \node[word]  (1s2) at (\ox+0*\dx,\sy+0*\dy) {};
    \node[word]  (1s3) at (\ox+0*\dx,\sy+1*\dy) {};
    \node[sbox] (1sb) at  (\ox+0*\dx,\sy+0*\dy) {};

    \node[comp] (1c) at  (\ox+0.5*\oxb,\sy-1.5*\dy) {composition};

    \node[tracker] (1t) at  (\ox+0*\dx,5*\dy) {tracking};
    \node[softmax] (1sm) at  (\ox+3*\dx,7*\dy) {$\sigma$};
    \node[result] (1so) at  (\ox+3*\dx,9*\dy) {\shift};

    \def\ox{2*\oxb}

    \node[focus1] (2bb) at  (\ox+0*\dx,2*\dy) {};
    \node[word]  (2b3) at (\ox+0*\dx,\by-1*\dy) {};
    \node[word]  (2b2) at (\ox+0*\dx,\by+0*\dy) {};
    \node[word]  (2b1) at (\ox+0*\dx,\by+1*\dy) {down};
    \node[bbox] (2bb) at  (\ox+0*\dx,\by+0.5*\dy) {};

    \node[focus2] (2sb) at  (\ox+0*\dx,\sy-0.5*\dy) {};
    \node[word]  (2s1) at (\ox+0*\dx,\sy-1*\dy) {sat};
    \node[word]  (2s2) at (\ox+0*\dx,\sy+0*\dy) {the cat};
    \node[word]  (2s3) at (\ox+0*\dx,\sy+1*\dy) {};
    \node[sbox] (2sb) at  (\ox+0*\dx,\sy+0*\dy) {};

    \node[tracker] (2t) at  (\ox+0*\dx,5*\dy) {tracking};

    \pgfsetarrowsend{latex}
    \tikzstyle{fwd} = [draw=black, line width=1pt]

    \draw [fwd] (0sb) -- (0t);
    \draw [fwd] (0bb) -- (0t);
    \draw [fwd] (0t) -- (0sm);
    \draw [fwd] (0sm) -- (0so);
    \draw [fwd] (0sb) -- (0c);

    \draw [fwd] (0t) -- (1t);
    \draw [fwd] (0sb) -- (1sb);
    \draw [fwd] (0bb) -- (1bb);
    \draw [fwd] (0so) -- (1sb);
    \draw [fwd] (0so) -- (1bb);
    \draw [fwd] (0bb) to[out=20,in=-110] (1sb);
    \draw [fwd] (0c) -- (1sb);

    \draw [fwd] (1sb) -- (1t);
    \draw [fwd] (1bb) -- (1t);
    \draw [fwd] (1t) -- (1sm);
    \draw [fwd] (1sm) -- (1so);
    \draw [fwd] (1sb) -- (1c);

    \draw [fwd] (1t) -- (2t);
    \draw [fwd] (1sb) -- (2sb);
    \draw [fwd] (1bb) -- (2bb);
    \draw [fwd] (1so) -- (2sb);
    \draw [fwd] (1so) -- (2bb);
    \draw [fwd] (1bb) to[out=20,in=-110] (2sb);
    \draw [fwd] (1c) -- (2sb);

    \draw [fwd] (2sb) -- (2t);
    \draw [fwd] (2bb) -- (2t);

  \end{tikzpicture}}

 \caption{The Model 1/2 network unrolled for two transitions on the input \word{the cat sat down}.}\label{fig:model:top}

\end{subfigure}\\\\\\
\begin{subfigure}[t]{\textwidth}
\centering
\scalebox{0.6}{
 \begin{tikzpicture}
    \def\dx{21pt}
    \def\dy{11pt}
    \def\sy{7*\dy}
    \def\oxb{5.5*\dx}
    \def\by{1*\dy}
    \def\ox{0*\oxb}

    \tikzstyle{word}=[text width=35mm,align=center,text height=2mm]    
    \tikzstyle{tracker}=[fill=red!40,text width=15mm,align=center,text height=2mm]
    \tikzstyle{softmax}=[fill=orange!40,text width=33mm,align=center,text height=2mm]
    \tikzstyle{comp}=[fill=green!40,text width=20mm,align=center,text height=2mm]
    \tikzstyle{result}=[line width=1pt,draw=black,text width=15mm,align=center,text height=2mm]    
    \tikzstyle{sbox}=[line width=1pt,draw=black,text width=33mm,align=center,text height=10mm]
    \tikzstyle{bbox}=[line width=1pt,draw=black,text width=33mm,align=center,text height=6.5mm]
    \tikzstyle{focus1}=[fill=yellow!80,opacity=0.5,text width=33mm,align=center,text height=2mm]
    \tikzstyle{focus2}=[fill=yellow!80,opacity=0.5,text width=33mm,align=center,text height=5.5mm]

    \def\ox{0*\oxb}

    \node[focus1] (1bb) at  (\ox+0*\dx,2*\dy) {};
    \node[word]  (1b3) at (\ox+0*\dx,\by-1*\dy) {};
    \node[word]  (1b2) at (\ox+0*\dx,\by+0*\dy) {down};
    \node[word]  (1b1) at (\ox+0*\dx,\by+1*\dy) {sat};
    \node[bbox] (1bb) at  (\ox+0*\dx,\by+0.5*\dy) {};

    \node[focus2] (1sb) at  (\ox+0*\dx,\sy-0.5*\dy) {};
    \node[word]  (1s1) at (\ox+0*\dx,\sy-1*\dy) {cat};
    \node[word]  (1s2) at (\ox+0*\dx,\sy+0*\dy) {the};
    \node[word]  (1s3) at (\ox+0*\dx,\sy+1*\dy) {};
    \node[sbox] (1sb) at  (\ox+0*\dx,\sy+0*\dy) {};

    \node[result] (1so) at  (\ox+0.5*\dx,4*\dy) {\reduce};

    \def\ox{1*\oxb}

    \node[focus1] (2bb) at  (\ox+0*\dx,2*\dy) {};
    \node[word]  (2b3) at (\ox+0*\dx,\by-1*\dy) {};
    \node[word]  (2b2) at (\ox+0*\dx,\by+0*\dy) {down};
    \node[word]  (2b1) at (\ox+0*\dx,\by+1*\dy) {sat};
    \node[bbox] (2bb) at  (\ox+0*\dx,\by+0.5*\dy) {};

    \node[focus2] (2sb) at  (\ox+0*\dx,\sy-0.5*\dy) {};
    \node[word]  (2s1) at (\ox+0*\dx,\sy-1*\dy) {the cat};
    \node[word]  (2s2) at (\ox+0*\dx,\sy+0*\dy) {};
    \node[word]  (2s3) at (\ox+0*\dx,\sy+1*\dy) {};
    \node[sbox] (2sb) at  (\ox+0*\dx,\sy+0*\dy) {};

    \node[result] (2so) at  (\ox+0.5*\dx,4*\dy) {\shift};

    \def\ox{2*\oxb}

    \node[focus1] (3bb) at  (\ox+0*\dx,2*\dy) {};
    \node[word]  (3b3) at (\ox+0*\dx,\by-1*\dy) {};
    \node[word]  (3b2) at (\ox+0*\dx,\by+0*\dy) {};
    \node[word]  (3b1) at (\ox+0*\dx,\by+1*\dy) {down};
    \node[bbox] (3bb) at  (\ox+0*\dx,\by+0.5*\dy) {};

    \node[focus2] (3sb) at  (\ox+0*\dx,\sy-0.5*\dy) {};
    \node[word]  (3s1) at (\ox+0*\dx,\sy-1*\dy) {sat};
    \node[word]  (3s2) at (\ox+0*\dx,\sy+0*\dy) {the cat};
    \node[word]  (3s3) at (\ox+0*\dx,\sy+1*\dy) {};
    \node[sbox] (3sb) at  (\ox+0*\dx,\sy+0*\dy) {};

    \node[result] (3so) at  (\ox+0.5*\dx,4*\dy) {\shift};

    \def\ox{3*\oxb}

    \node[focus1] (4bb) at  (\ox+0*\dx,2*\dy) {};
    \node[word]  (4b3) at (\ox+0*\dx,\by-1*\dy) {};
    \node[word]  (4b2) at (\ox+0*\dx,\by+0*\dy) {};
    \node[word]  (4b1) at (\ox+0*\dx,\by+1*\dy) {};
    \node[bbox] (4bb) at  (\ox+0*\dx,\by+0.5*\dy) {};

    \node[focus2] (4sb) at  (\ox+0*\dx,\sy-0.5*\dy) {};
    \node[word]  (4s1) at (\ox+0*\dx,\sy-1*\dy) {down};
    \node[word]  (4s2) at (\ox+0*\dx,\sy+0*\dy) {sat};
    \node[word]  (4s3) at (\ox+0*\dx,\sy+1*\dy) {the cat};
    \node[sbox] (4sb) at  (\ox+0*\dx,\sy+0*\dy) {};

    \node[result] (4so) at  (\ox+0.5*\dx,4*\dy) {\reduce};

    \def\ox{4*\oxb}

    \node[focus1] (5bb) at  (\ox+0*\dx,2*\dy) {};
    \node[word]  (5b3) at (\ox+0*\dx,\by-1*\dy) {};
    \node[word]  (5b2) at (\ox+0*\dx,\by+0*\dy) {};
    \node[word]  (5b1) at (\ox+0*\dx,\by+1*\dy) {};
    \node[bbox] (5bb) at  (\ox+0*\dx,\by+0.5*\dy) {};

    \node[focus2] (5sb) at  (\ox+0*\dx,\sy-0.5*\dy) {};
    \node[word]  (5s1) at (\ox+0*\dx,\sy-1*\dy) {sat down};
    \node[word]  (5s2) at (\ox+0*\dx,\sy+0*\dy) {the cat};
    \node[word]  (5s3) at (\ox+0*\dx,\sy+1*\dy) {};
    \node[sbox] (5sb) at  (\ox+0*\dx,\sy+0*\dy) {};

    \node[result] (5so) at  (\ox+0.5*\dx,4*\dy) {\reduce};

    \def\ox{5*\oxb}

    \node[focus2] (6sb) at  (\ox+0*\dx,\sy-0.5*\dy) {};
    \node[word]  (6s1) at (\ox+0*\dx,\sy-1*\dy) {(the cat) (sat down)};
    \node[word]  (6s2) at (\ox+0*\dx,\sy+0*\dy) {};
    \node[word]  (6s3) at (\ox+0*\dx,\sy+1*\dy) {};
    \node[sbox] (6sb) at  (\ox+0*\dx,\sy+0*\dy) {};

    \node[softmax] (6sm) at  (\ox+0.5*\dx,2*\dy) {output to semantic task model};

    \pgfsetarrowsend{latex}
    \tikzstyle{fwd} = [draw=black, line width=1pt]

   \draw [fwd] (1sb) -- (1so);
   \draw [fwd] (1bb) -- (1so);

    \draw [fwd] (1sb) -- (2sb);
    \draw [fwd] (1bb) -- (2bb);
    \draw [fwd] (1so) -- (2sb);
    \draw [fwd] (1so) -- (2bb);
    \draw [fwd] (1bb) -- (2sb);

   \draw [fwd] (2sb) -- (2so);
   \draw [fwd] (2bb) -- (2so);

    \draw [fwd] (2sb) -- (3sb);
    \draw [fwd] (2bb) -- (3bb);
    \draw [fwd] (2so) -- (3sb);
    \draw [fwd] (2so) -- (3bb);
    \draw [fwd] (2bb) -- (3sb);

   \draw [fwd] (3sb) -- (3so);
   \draw [fwd] (3bb) -- (3so);

    \draw [fwd] (3sb) -- (4sb);
    \draw [fwd] (3bb) -- (4bb);
    \draw [fwd] (3so) -- (4sb);
    \draw [fwd] (3so) -- (4bb);
    \draw [fwd] (3bb) -- (4sb);

   \draw [fwd] (4sb) -- (4so);
   \draw [fwd] (4bb) -- (4so);

    \draw [fwd] (4sb) -- (5sb);
    \draw [fwd] (4bb) -- (5bb);
    \draw [fwd] (4so) -- (5sb);
    \draw [fwd] (4so) -- (5bb);
    \draw [fwd] (4bb) -- (5sb);

   \draw [fwd] (5sb) -- (5so);
   \draw [fwd] (5bb) -- (5so);

    \draw [fwd] (5sb) -- (6sb);
    \draw [fwd] (5so) -- (6sb);

   \draw [fwd] (6sb) -- (6sm);

  \end{tikzpicture}}

 \caption{The fully unrolled Model 1/2 network for \word{the cat sat down} with some layers omitted for clarity.}\label{fig:model:top}  
\end{subfigure}
\caption{\label{m1-views}Two views of Models 1 and 2 (which use equivalent model graphs). In both views, the lower boxes represent the input buffer, and the upper boxes represent the stack. Yellow highlighting indicates which portions of these data structures are visible to the tracking LSTM and to the composition function.}
\end{figure*}