### 2016-02-24 writing/tree_attn_fig.tex
Initial outline of the paper.

%!TEX root = hard_stack_paper/paper.tex

\begin{figure}[t]
\centering
\scalebox{0.9}{
 \begin{tikzpicture}
    \tikzstyle{word}=[fill=yellow!40,text height=2mm]    
    \tikzstyle{nonleaf}=[fill=yellow!40,text height=2mm]

    \begin{scope}[shift={(0in,0in)}, frontier/.style={distance from root=60pt}]

    \Tree [.\node[nonleaf](1thecatsatdown){the cat sat down}; [.\node[nonleaf](1thecat){the cat}; \node[word](1the){the}; \node[word](1cat){cat}; ] [.\node[nonleaf](1satdown){sat down}; \node[word](1sat){sat}; \node[word](1down){down}; ] ]

    \end{scope}

    \begin{scope}[shift={(2in,0in)}, frontier/.style={distance from root=60pt}]

    \Tree [.\node[nonleaf](2thekittenmeowed){the kitten meowed}; [.\node[nonleaf](2thekitten){the kitten}; \node[word](2the){the}; \node[word](2kitten){kitten}; ] \node[word](2meowed){meowed}; ]

    \end{scope}

    \usetikzlibrary{arrows,scopes}

    \tikzstyle{attn} = [line width=1pt,draw=black!80,opacity=0.25,line cap=round]
    \tikzstyle{heavy} = [line width=3pt]
    \tikzstyle{light} = [line width=0.33pt]
    \tikzstyle{focus} = [draw=red,opacity=1.0]

    \draw [attn,heavy] (1the) to[in=155,out=25] (2the);
    \draw [attn,light] (1thecat) to[in=155,out=25] (2the);

    \draw [attn,heavy] (1cat) to[in=155,out=25] (2kitten);
    \draw [attn,light] (1thecat) to[in=155,out=25] (2kitten);
    \draw [attn,light] (1thecatsatdown) to[in=155,out=25] (2kitten);

    \draw [attn,light] (1satdown) to[in=155,out=25] (2thekittenmeowed);
    \draw [attn,light] (1thecat) to[in=155,out=25] (2thekittenmeowed);
    \draw [attn,heavy] (1thecatsatdown) to[in=155,out=25] (2thekittenmeowed);

    \draw [attn] (1satdown) to[in=155,out=25] (2meowed);
    \draw [attn] (1sat) to[in=155,out=25] (2meowed);
    \draw [attn,light] (1down) to[in=155,out=25] (2meowed);

    \draw [attn,focus] (1cat) to[in=155,out=25,] (2thekitten);
    \draw [attn,heavy,focus] (1thecat) to[in=155,out=25] (2thekitten);
    \draw [attn,light,focus] (1the) to[in=155,out=25] (2thekitten);
    \draw [attn,light,focus] (1thecatsatdown) to[in=155,out=25] (2thekitten);

  \end{tikzpicture}}

 \caption{Soft alignments between the nodes of two trees, with alignments to ``the kitten'' highlighted in red.}\label{fig:tree_attn}
\end{figure}